---
title: "Untitled"
author: "Tobi Bosede"
date: "April 4, 2017"
header-includes:
- \usepackage{multirow}

output: 
  pdf_document:
  #variant: markdown+raw_tex (alternate to below)
  keep_tex: yes
  latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
\begin{tabular}{|l|l|}\hline
Age & Frequency \\ \hline
18--25  & 15 \\
26--35  & 33 \\
36--45  & 22 \\ \hline
\end{tabular}


\begin{table}
\centering
\caption{}
\label{table-paramvalues}
\begin{tabular}{ p{4cm} p{4cm} p{4cm} }
\hline \\ [-1.5ex]
colname &amp; colname &amp; colname \\ [1ex]
\hline \\ [-1.5ex]
Info & info & info \\ [1ex]
Info & info & info \\ [1ex]
Info & info & info \\ [1ex]
\hline
\end{tabular}
\end{table}




```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
