---
title: "Trade Count Regression"
author: "Tobi Bosede"
date: "March 22, 2017"
output: html_document
---
550.700
```{r, message=FALSE, warning=FALSE}
library(mgcv)
library(ggplot2)
```

```{r}
setwd("C:/Users/aniverb/Documents/Grad_School/JHU/700")
trade_data=read.table("trades_count_regression_2016-11-18.txt", sep="\t", header=T)
trade_data=trade_data[order(trade_data$Date),]
head(trade_data)
```

```{r}
#use only half of data randomly split into train and test
n=nrow(trade_data)
lastDate=as.vector(trade_data$Date[n/2])
nRem=nrow(subset(trade_data[(n/2+1):n,], Date==lastDate))
train=trade_data[1:(n/2+nRem), ] #past data, ends with 8/2 data 
set.seed(12-1-16)
test_id=sample((n/2+nRem+1):n, (n/2-nRem))[1:floor((n/2-nRem)/2)]
test1=trade_data[test_id, ] #future data
test2=trade_data[-test_id, ] #reserve data
```

Build linear regression model with penalized cubic spline. Spline parameters found by cross validation.
```{r}
model1 <- gam(TradeTotBCTrans ~ s(TimeToMaturity, bs="cr") + s(DayofMonth, bs="cr") + factor(ProductType) + factor(ProductName) + factor(DayOfTheWeek), family=gaussian(), data = train, method="GCV.Cp", select=TRUE)
summary(model1)
```
The model is not very good as the Adjust R sq is barely .5 as well, only about half of the deviance/variation is explained.

```{r}
plot(model1)
p <- predict(model1, type="lpmatrix")

```{r}
beta <- coef(model1)[grepl("TimeToMaturity", names(coef(model1)))]
s <- p[,grepl("TimeToMaturity", colnames(p))] %*% beta
ggplot(data=cbind.data.frame(s, train$TimeToMaturity), aes(x=train$TimeToMaturity, y=s, ymin = -5, ymax = 3)) + geom_line()
beta <- coef(model1)[grepl("DayofMonth", names(coef(model1)))]
s <- p[,grepl("DayofMonth", colnames(p))] %*% beta
ggplot(data=cbind.data.frame(s, train$DayofMonth), aes(x=train$DayofMonth, y=s, ymin = -5, ymax = 3)) + geom_line()
```

```{r}
model1$coefficients
model1$sp #spline parameters
```

Unsurprisingly, the below residuals are quite large, especially since they are for the box cox transformed trade counts. However they seem constant so an assumption of the linear regression model is fufilled.
```{r}
res=model1$residuals
plot(res, pch=20)
hist(res)
```

Actual values are black, predicted values are red
```{r}
trainPred=model1$fitted.values
plot(train$TimeToMaturity, train$TradeTotBCTrans, pch=20, ylim=c(min(train$TradeTotBCTrans, trainPred), max(train$TradeTotBCTrans, trainPred)))
points(train$TimeToMaturity, trainPred, col="red", pch=20)
legend("topright", c("actual", "predicted"),lwd=c(3,3),col=c("black", "red"))

plot(train$Day, train$TradeTotBCTrans, pch=20, ylim=c(min(train$TradeTotBCTrans, trainPred), max(train$TradeTotBCTrans, trainPred)))
points(train$Day, trainPred, col="red", pch=20)
```

Model trained on 1/2 of past data so only testing on 1/4 of randomized future data.
```{r}
new_data1=test1
head(new_data1)
new_data1Pred <- predict(model1, new_data1)
new_data1["Prediction"]=new_data1Pred
plot(new_data1$TimeToMaturity, new_data1$TradeTotBCTrans, pch=20, ylim=c(min(new_data1$TradeTotBCTrans, new_data1Pred), max(new_data1$TradeTotBCTrans, new_data1Pred)))
points(new_data1$TimeToMaturity, new_data1$Prediction, col="red", pch=20)
legend("topright", c("actual", "predicted"),lwd=c(3,3),col=c("black", "red"))

plot(new_data1$Day, new_data1$TradeTotBCTrans, pch=20, ylim=c(min(new_data1$TradeTotBCTrans, new_data1Pred), max(new_data1$TradeTotBCTrans, new_data1Pred)))
points(new_data1$Day, new_data1$Prediction, col="red", pch=20)
```

```{r}
model1prime <- gam(TradeTotBCTrans ~ s(TimeToMaturity, bs = "cr") + s(DayofMonth, bs = "cr") + factor(ProductType) +   factor(DayOfTheWeek), family=gaussian(), data = train, method="GCV.Cp", select=TRUE)
summary(model1prime)
new_data1PredP <- predict(model1prime, new_data1)
new_data1["PredictionPrime"]=new_data1PredP
```

Calculating mean absolute error (MAE) and means for each product type.
```{r}
types=c('Metal', 'Bond', 'Agr', 'Energy', 'Index', 'FX')
MAEone_mod=c()
MAEone_modP=c()
medians=c()
count=0
for (type in types){
  count=count+1
  typsub=subset(new_data1, ProductType==type)
  MAEone_mod[count]= mean(abs(typsub$TradeTotBCTrans-typsub$Prediction))
  MAEone_modP[count]= mean(abs(typsub$TradeTotBCTrans-typsub$PredictionPrime))
  medians[count]=median(typsub$TradeTotBCTrans)
}
```

Creating models for each product type.
```{r}
placeH=data.frame(Instrument=NULL, STTM=NULL, TimeToMat=NULL)
placeHD=data.frame(Instrument=NULL, sDoM=NULL, DoM=NULL)
MAEmult_mod=c()
MAEmult_modPrime=c()
all_products=c()
all_medians=c()
all_MAE=c()
all_MAEprime=c()
count=0
for (type in types){
  count=count+1
  typsub=subset(train, ProductType==type)
  #model2 
  model <- gam(TradeTotBCTrans ~ s(TimeToMaturity, bs="cr") + s(DayofMonth, bs="cr") + factor(ProductName) + factor(DayOfTheWeek), family=gaussian(), data = typsub, method="GCV.Cp", select=TRUE)
  print(summary(model))
  plot(model)
  
  #superimposed spline plots
 p <- predict(model, type="lpmatrix")
 
 beta <- coef(model)[grepl("TimeToMaturity", names(coef(model)))]
 sTTM <- p[,grepl("TimeToMaturity", colnames(p))] %*% beta
 placeH= rbind(placeH, data.frame(Instrument=rep(type, nrow(typsub)), STTM=sTTM, TTM=typsub$TimeToMaturity))
  
 beta <- coef(model)[grepl("DayofMonth", names(coef(model)))]
 sDoM <- p[,grepl("DayofMonth", colnames(p))] %*% beta
 placeHD = rbind(placeHD, data.frame(Instrument=rep(type, nrow(typsub)), sDoM=sDoM, DoM=typsub$DayofMonth))
  
 #Model2'
  modelPrime <- gam(TradeTotBCTrans ~ s(TimeToMaturity, bs = "cr") + s(DayofMonth, bs = "cr") +  factor(DayOfTheWeek), family=gaussian(), data = typsub, method="GCV.Cp", select=TRUE)
  print(summary(modelPrime))
  plot(modelPrime)
  
  typsub_new=subset(new_data1, ProductType==type)
  new_data1Pred <- predict(model, typsub_new)
  typsub_new["Prediction"]=new_data1Pred
  cat(type)
  plot(typsub_new$TimeToMaturity, typsub_new$TradeTotBCTrans, pch=20, ylim=c(min(typsub_new$TradeTotBCTrans, typsub_new$Prediction), max(typsub_new$TradeTotBCTrans, typsub_new$Prediction)))
  points(typsub_new$TimeToMaturity, typsub_new$Prediction, col="red", pch=20)
  legend("topright", c("actual", "predicted"),lwd=c(3,3),col=c("black", "red"))
  plot(typsub_new$Day, typsub_new$TradeTotBCTrans, pch=20, ylim=c(min(typsub_new$TradeTotBCTrans, typsub_new$Prediction), max(typsub_new$TradeTotBCTrans, typsub_new$Prediction)))
  points(typsub_new$Day, typsub_new$Prediction, col="red", pch=20)
  MAEmult_mod[count]= mean(abs(typsub_new$TradeTotBCTrans-typsub_new$Prediction))
  
  typsub_newT=typsub_new[rep(1,2200),]
  true=typsub_newT$Day
  set.seed(1-13-17)
  typsub_newT['Day']=sample(rep(1:140, 16)[1:2200], 2200)
  print(head(typsub_newT))
  new_data1PredT <- predict(model, typsub_newT)
  typsub_newT["Prediction"]=new_data1PredT
  
  new_data1PredTP <- predict(modelPrime, typsub_newT)
  typsub_newT["PredictionPrime"]=new_data1PredTP
  
  legend("topright", c("actual", "predicted"), lwd=c(3,3), col=c("black", "red"))
  plot(typsub_newT$Day, typsub_newT$TradeTotBCTrans, pch=20, ylim=c(min(typsub_newT$TradeTotBCTrans, typsub_newT$Prediction), max(typsub_newT$TradeTotBCTrans, typsub_newT$Prediction)))
  points(typsub_newT$Day, typsub_newT$Prediction, col="red", pch=20)
  cat("True Day:", true[1], "\n", "\n")
  
  MAEmult_modP=c()
  MAEmult_modPprime=c()
  medianP=c()
  countP=0
  products=as.vector(unique(typsub_new$ProductName))
  all_products=append(all_products,products,length(all_products))
  for (product in products){
      countP=countP+1
      prodsub=subset(typsub_new, ProductName==product)
      MAEmult_modP[countP]= mean(abs(prodsub$TradeTotBCTrans-prodsub$Prediction))
      MAEmult_modPprime[countP]= mean(abs(prodsub$TradeTotBCTrans-prodsub$PredictionPrime))
      medianP[countP]=median(prodsub$TradeTotBCTrans)
  }
  all_MAE=append(all_MAE,MAEmult_modP,length(all_MAE))
  all_MAEprime=append(all_MAEprime,MAEmult_modPprime,length(all_MAE))
  all_medians=append(all_medians, medianP, length(all_medians))
}
p<-ggplot(placeH, aes(x=TTM, y=STTM, group=Instrument, ymin = -10, ymax = 10)) + geom_line(aes(color=Instrument), size=1, alpha = 0.6)+ ylab("Trade Total Box Cox Trans")+xlab("Time to Maturity") + labs(title="Model 2 Spline on Time to Maturity")+ theme_classic()+scale_color_brewer(palette="Dark2") 
p

pD<-ggplot(placeHD, aes(x=DoM, y=sDoM, group=Instrument, ymin = -10, ymax = 10)) + geom_line(aes(color=Instrument), size=1, alpha = 0.6)+ ylab("Trade Total Box Cox Trans")+xlab("Day of Month") + labs(title="Model 2 Spline on Day of the Month")+ theme_classic()+scale_color_brewer(palette="Dark2") 
pD

data.frame(types, Model2=MAEmult_mod, Model2Prime=MAEmult_modPprime, Model1=MAEone_mod, Model1Prime=MAEone_modP, Median=medians) #separate models for each product type
#comparing all 4 models

data.frame(all_products, Model2=all_MAE, Model2Prime=all_MAEprime, Median=all_medians) #comparing just separate models
```
Removing spline from day improves future prediction. In general the huge variances in the medians of products as opposed to type seems to indicate that we should drill down even further and make multiple models for each product .
